First give a general overview of

% !TEX root = ../main.tex
\chapter{Derivation of the Redfield Equation} % Main chapter title
\label{Chapter:Derivation_Redfield_Equation} % Label for referencing this chapter

The following derivation is part of \cite{Manzano2020ShortIntroductionLindblad}
%------------------------------------------------------------------------------
%	SECTION 1
%------------------------------------------------------------------------------

\section{Derivation from microscopic dynamics}
\label{sec:Derivation_redfield_eq_from_microscopic_dynamics}
The most common derivation of the Redfield master equation is based on open quantum theory.
We will begin by discussing the general problem, where a small quantum system interacts with a larger environment.
The total system Hilbert space $\mathcal{H}_T$ is divided into our system of interest, belonging to a Hilbert space $\mathcal{H}_S$, and the environment living in $\mathcal{H}_E$.

\begin{equation}
    \mathcal{H}_T = \mathcal{H}_S \otimes \mathcal{H}_E.
    \label{eq:Total_Hilbert_Space}
\end{equation}
The Redfield equation is then an effective motion equation for a this subsystem of interest $ S $.
The derivation can be found in several textbooks such as Breuer and Petruccione \cite{BreuerPetruccione2009TheoryOpenQuantum}.

The evolution of the total system is given by the Liouville-Von Neumann equation, which represents the starting point of our derivation:

\begin{equation}
    \dot{\rho}_T(t) = -i[H_T, \rho_T(t)],
    \label{eq:Von_Neumann_Equation}
\end{equation}
where $\rho_T(t)$ is the density matrix of the total system, and $H_T$ is the total Hamiltonian.
Note that we used units where $\hbar = 1$.
This equation can not be solved for arbitrary large enviroments.
But as we are interested in the dynamics of the system without the environment,
we trace over the environment degrees of freedom to obtain the reduced density matrix of the system
$\rho(t) = \mathrm{Tr}_E[\rho_T]$.

The total Hamiltonian can be separated as:

\begin{equation}
    H_T = H_S \otimes \mathbb{1}_E + \mathbb{1}_S \otimes H_E + \alpha H_I,
    \label{eq:Total_Hamiltonian}
\end{equation}
where $H_S \in \mathcal{H}_S$, $H_E \in \mathcal{H}_E$, and $H_I \in \mathcal{H}_T$. $ H_I $ represents the interaction between the system and the environment with coupling strength $\alpha$.
The interaction term is typically decomposed as:

\begin{equation}
    H_I = \sum_i S_i \otimes E_i,
    \label{eq:Interaction_Hamiltonian}
\end{equation}
where $S_i \in \mathcal{B}(\mathcal{H}_S)$ and $E_i \in \mathcal{B}(\mathcal{H}_E)$ are operators, that only act on the system and environment respectively.

The following requirements are to be fuildilled by the final derived Redfield equation:

\begin{itemize}
    \item The equation should be linear in the system density matrix $ \dot{ \rho}_S(t) = F(\rho_S(t))$ (reduced equation of motion).
    \item The equation should be Markovian, meaning that the evolution of the system density matrix at time $t$ only depends on the state of the system at time $t$ and not on its past history $ \dot{ \rho}_S(t) = \rho_S(t)$ .
    \item The equation should be trace-preserving, meaning that $\mathrm{Tr}[\rho_S(t)] = \mathrm{Tr}[\rho_S(0)]$ for all times $t$.
\end{itemize}
Unlike the Linblad equation it does not guarantee the complete (not even normal) positivity of the density matrix, which is a requirement for a physical state.
So care has to be taken, when the Redfieldequation is useful.
The equation will be valid in the weak coupling limit, meaning that the constant in the interaction Hamiltonian $H_I$ fulfills  $\alpha \ll  1$.

%----------------------------------------------------------------------------------------
%	SUBSECTION 1
%----------------------------------------------------------------------------------------

\subsection{Interaction Picture}
\label{subsec:Interaction_Picture}

To describe the system dynamics, we move to the interaction picture where the operators evolve with respect to $H_S + H_E$.
Any arbitrary operator $O$ in the Schrödinger picture takes the form

\begin{equation}
    \hat{O}(t) = e^{i(H_S+H_E)t} O e^{-i(H_S+H_E)t},
    \label{eq:Interaction_Picture_Operators}
\end{equation}
in the interaction picture, and depends on time.
States now evolve only according to the interaction Hamiltonian $H_I$, and the Liouville-Von Neumann equation (Eq. \eqref{eq:Von_Neumann_Equation}) changes to:

\begin{equation}
    \dot{\hat{\rho}}_T(t) = -i \alpha [\hat{H}_I(t), \hat{\rho}_T(t)],
    \label{eq:LiouvilleVN}
\end{equation}
which can be formally integrated as:

\begin{equation}
    \hat{\rho}_T(t) = \hat{\rho}_T(0) - i \alpha \int_0^t ds [\hat{H}_I(s), \hat{\rho}_T(s)].
    \label{eq:Formal_Integration}
\end{equation}
This equation will be inserted in Eq. \eqref{eq:LiouvilleVN}:

\begin{equation}
    \dot{\hat{\rho}}_T(t) = -i \alpha \left[ \hat{H}_I(t), \hat{\rho}_T(0) \right]
    - \alpha^2 \int_0^t \left[ \hat{H}_I(t), \left[ \hat{H}_I(s), \hat{\rho}_T(s) \right] \right] ds,
    \label{eq:Second_Order_Expansion}
\end{equation}
which is not Morkovian, because of the intergal which sums up all the past.
The step of integration and insertion can be repreated leading to a series expansion of the density matrix in the small parameter $\alpha$:

\begin{equation}
    \dot{\hat{\rho}}_T(t) = -i \alpha \underbrace{\left[ \hat{H}_I(t), \hat{\rho}_T(0) \right]}_{\text{(1)}}
    - \alpha^2 \int_0^t \underbrace{\left[ \hat{H}_I(t), \left[ \hat{H}_I(s), \hat{\rho}_T(t) \right] \right]}_{\text{(2)}} ds + \mathcal{O} (\alpha^3).
    \label{eq:Second_Order_Expansion_wo_third}
\end{equation}
where third order contributions (and higher) are neglected from now on.
This can be justified in the weak coupling limit where $ \alpha \ll 1 $, which represents the "\textbf{Born}" approximation.

In this sense, the Redfield equation will be a second-order approximation of the actual dynamics.
The weak coupling assumption does not hold universally and cannot be applied to all systems.
For instance, it is often invalid in chemical or biological systems.

Remark, that Eq. \eqref{eq:Second_Order_Expansion_wo_third} is still not Morkovian, because of the intergal over time.
Since we are only interested in the dynamics of the system $ S $, we will now take the partial trace over the environment degrees of freedom in Eq. \eqref{eq:Second_Order_Expansion_wo_third}.

%----------------------------------------------------------------------------------------
%	SUBSECTION 2
%----------------------------------------------------------------------------------------

\subsection{Partial Trace}
\label{subsec:Partial_Trace}

We now assume the whole system to be seperatable at $ t = 0 $ as a product state:

\begin{equation}
    \hat{\rho}_T(0) = \hat{\rho}_S(0) \otimes \hat{\rho}_E(0),
    \label{eq:Initial_Product_State}
\end{equation}
which means, that the two sub-systems only come into contact at $ t = 0 $ and there are no correlations.
We take the interaction Hamiltonian Eq. \eqref{eq:Interaction_Hamiltonian}, which has the same shape in the Interaction Picture into account.
With this, the partial trace over the environment of the part (1) of Eq. \eqref{eq:Second_Order_Expansion_wo_third} is given by:

\begin{align}
    \sum_i \mathrm{Tr}_E\big[ S_i \otimes E_i , \hat{\rho}_S(0) \otimes \hat{\rho}_E(0)\big]
    = \sum_i \big(S_i \hat{\rho}_S(0) - \hat{\rho}_S(0) S_i\big) \cdot \mathrm{Tr}_E \big[E_i \hat{\rho}_E(0)\big],
    \label{eq:Trace_Relation_first_part}
\end{align}

where we used the cyclic property of the trace.
We define the average of the bath degrees of freedom at zero temperature:

\begin{equation}
    \langle E_i \rangle_0 \equiv \mathrm{Tr}_E \big[E_i \hat{\rho}_E(0)\big].
    \label{eq:Environment_Expectation_Value}
\end{equation}
which results to zero, simplifying Eq. \eqref{eq:Second_Order_Expansion_wo_third} to only the second part.
This can always be justified when adding a zero to the total Hamiltonian:

\begin{equation}
    H_T = H_S' + H_E + \alpha H_I',
    \label{eq:Shifted_Total_Hamiltonian}
\end{equation}
where a new interaction and system Hamiltonian are introduced.
The interaction Hamiltonian takes new environmental operators $E_i'$ which are shifted by the average of the environment operators at time $t = 0$:

\begin{equation}
    H_I' = \sum_i S_i \otimes E_i' = \sum_i S_i \otimes (E_i - \langle E_i \rangle_0).
    \label{eq:Shifted_Interaction_Hamiltonian}
\end{equation}
The new system Hamiltonian is then given by the sum of the original system Hamiltonian shifted proportionally by the average of the environment operators at time $t = 0$:

\begin{equation}
    H_S' = H_S + \alpha \sum_i S_i \langle E_i \rangle_0,
    \label{eq:Shifted_System_Hamiltonian}
\end{equation}
This however doesn't change the structure of the system dynamics.
It only accounts for a redefinition of the energy levels ("a sort of renormalization").
This way the equation \eqref{eq:Trace_Relation_first_part} results to zero and only the second part of the equation Eq. \eqref{eq:Second_Order_Expansion_wo_third} remains.

\begin{align}
    \dot{\rho}_S(t) = -i \alpha [\hat{H}_I(t),\hat{\rho}_T(0)]
      & - \alpha^2 \int_0^t ds \mathrm{Tr}_E \big[\hat{H}_I(t), [\hat{H}_I(s), \rho_S(t) \otimes \rho_E]\big] \notag \\
    = & - \alpha^2 \int_0^t ds \mathrm{Tr}_E \big[\hat{H}_I(t), [\hat{H}_I(s), \rho_S(t) \otimes \rho_E]\big].
    \label{eq:Partial_Trace_Derivation}
\end{align}
In the follwoing, we will derive the final expression by calculating the environmental traces in the last equation.

%----------------------------------------------------------------------------------------
%	SUBSECTION 3
%----------------------------------------------------------------------------------------

\subsection{Final Expression}
\label{subsec:Final_Expression}

Defining $s' = t - s$, we rewrite the second-order term as:
\begin{align}
    \dot{\rho}_S(t)  = \alpha^2 \int_0^t ds \mathrm{Tr}_E \bigg\{
    \hat{H}_I(t) \big[ \hat{H}_I(t-s) \hat{\rho}_T(t) - \hat{\rho}_T(t) \hat{H}_I(t-s) \big] \notag \\
    - \big[ \hat{H}_I(t-s) \hat{\rho}_T(t) - \hat{\rho}_T(t) \hat{H}_I(t-s) \big] \hat{H}_I(t)
    \bigg\}.
    \label{eq:Second_Order_Final_Expression}
\end{align}
A seperatability at all times is now assumed:

\begin{equation}
    \hat{\rho}_T(t) = \hat{\rho}_S(t) \otimes \hat{\rho}_E(t),
    \label{eq:Reduced_Density_Matrix_Assumption}
\end{equation}
This assumtion has to be made even stronger later $\hat{\rho}_T(t) = \hat{\rho}_S(t) \otimes \hat{\rho}_E(0)$.
Expanding Eq. \eqref{eq:Second_Order_Final_Expression}, we obtain:

\begin{align}
    \dot{\hat{\rho}}_T(t) & =  \alpha^2 \int_0^t ds
    \bigg\{
    \mathrm{Tr}_E \big[ \hat{H}_I(t) \hat{H}_I(t-s) \hat{\rho}_T(t) \big] -
    \mathrm{Tr}_E \big[ \hat{H}_I(t) \hat{\rho}_T(t) \hat{H}_I(t-s) \big] - \notag \\
                          & \qquad \qquad \qquad
    \mathrm{Tr}_E \big[ \hat{H}_I(t-s) \hat{\rho}_T(t) \hat{H}_I(t) \big] +
    \mathrm{Tr}_E \big[ \hat{\rho}_T(t) \hat{H}_I(t-s) \hat{H}_I(t) \big]
    \bigg\}.
    \label{eq:Expanded_Second_Order_Expression}
\end{align}

Now, inserting the interaction Hamiltonian by tracking the operators at time $t - s$ with $i'$ and at $t$ with $i$, we have:

\begin{align}
    \dot{\hat{\rho}}_T(t) & = \alpha^2  \sum_{i, i'} \int_0^t ds
    \bigg\{
    \mathrm{Tr}_E \big[ \hat{S}_i(t) \hat{S}_{i'}(t-s) \hat{\rho}_S(t)      \otimes   \hat{E}_{i}(t) \hat{E}_{i'}(t-s) \hat{\rho}_E(t)  \big] -  \notag                         \\
                          & \mathrm{Tr}_E \big[ \hat{S}_i(t) \hat{\rho}_S(t) \hat{S}_{i'}(t-s)      \otimes   \hat{E}_{i}(t) \hat{\rho}_E(t) \hat{E}_{i'}(t-s)  \big] - \notag  \\
                          & \mathrm{Tr}_E \big[ \hat{S}_{i'}(t-s) \hat{\rho}_S(t) \hat{S}_i(t)      \otimes   \hat{E}_{i'}(t-s) \hat{\rho}_E(t) \hat{E}_{i}(t)  \big] +  \notag \\
                          & \mathrm{Tr}_E \big[ \hat{\rho}_S(t) \hat{S}_{i'}(t-s) \hat{S}_i(t)      \otimes   \hat{\rho}_E(t) \hat{E}_{i'}(t-s) \hat{E}_{i}(t)  \big]
    \bigg\}.
    \label{eq:Interaction_Hamiltonian_Expansion}
\end{align}

Since the trace only acts on the environment, the system operators can be taken out of the trace, and we define the correlation functions:
\begin{equation}
    C_{ij}(t - s) = \mathrm{Tr}_E \big[\hat{E}_{i}(t) \hat{E}_{i'}(t-s) \hat{\rho}_E(t)\big],
    \label{eq:Environment_Correlation_Function}
\end{equation}
such that:
\begin{align}
    \dot{\hat{\rho}}_T(t) = \alpha^2  \sum_{i, i'} \int_0^t ds
    \bigg\{
    C_{ij}(t - s) \big[ \hat{S}_i(t),  \hat{S}_{i'}(t-s) \hat{\rho}_S(t) \big] + \text{H.c.}
    \bigg\}.
    \label{eq:Redfield_Equation_Final}
\end{align}
which is the desired form of the Redfield equation.

Note however, that the we have not used the strong condition $\hat{\rho}_T(t) = \hat{\rho}_S(t) \otimes \hat{\rho}_E(0)$ was not used yet.
This however will make it possible to calculate the correlations Eq. \eqref{eq:Environment_Correlation_Function}, because we can assume that the enviroment is in a thermal equilibruium at a certain temperature.
Because of the assumption this is the case for all times.
It is equivalent to say that the environment is unaffected by the system. It is memoryless, because it is very big.

%----------------------------------------------------------------------------------------
%	SECTION 3: PHYSICAL INTERPRETATION AND DETAILED BALANCE
%----------------------------------------------------------------------------------------

\section{Physical Interpretation: Emission and Absorption Processes}
\label{sec:emission_absorption_kms}

\noindent A fundamental question when applying the Bloch-Redfield formalism concerns whether both emission and absorption processes are correctly included in the theoretical description. This is particularly important when studying spectroscopic phenomena, where the competition between these processes determines the observed signals and the approach to thermal equilibrium.

\noindent The answer to this question is \textbf{Yes}, both spontaneous and stimulated emission, as well as thermal and induced absorption, are naturally included through the proper choice of the spectral function $S(\omega)$ that characterizes the system-environment interaction, provided this function satisfies the Kubo-Martin-Schwinger (KMS) condition for detailed balance \cite{MartinSchwinger1959TheoryManyparticleSystems, Sewell2020QuantumMechanicsIts}.

\subsection{The Kubo-Martin-Schwinger Condition}
\label{subsec:kms_condition}

\noindent The KMS condition requires that the spectral function satisfies the following relationship:

\begin{equation}
    S(-\omega) = e^{-\hbar\omega/(k_B T)} S(\omega)
    \label{eq:kms_condition}
\end{equation}

\noindent where $\hbar$ is the reduced Planck constant, $k_B$ is Boltzmann's constant, and $T$ is the temperature of the thermal environment. This condition ensures several crucial physical requirements:

\begin{itemize}
    \item \textbf{Emission processes:} Transitions with $\omega > 0$ correspond to energy transfer from the system to the environment, representing photon emission or energy dissipation.

    \item \textbf{Absorption processes:} Transitions with $\omega < 0$ correspond to energy transfer from the environment to the system, representing thermal excitation or energy absorption from the bath.

    \item \textbf{Detailed balance:} The ratio of forward and reverse transition rates follows the Boltzmann distribution, ensuring that the system approaches thermal equilibrium at temperature $T$ in the long-time limit.
\end{itemize}

\subsection{Physical Implications}
\label{subsec:physical_implications}

\noindent This microscopic formulation naturally accounts for the temperature dependence of physical observables. At low temperatures ($k_B T \ll \hbar\omega$), the exponential factor $e^{-\hbar\omega/(k_B T)} \approx 0$ for positive frequencies, meaning absorption processes are suppressed while emission dominates. Conversely, at high temperatures ($k_B T \gg \hbar\omega$), the exponential factor approaches unity, and both emission and absorption processes become equally probable.

\noindent The KMS condition is not merely a mathematical convenience but reflects the fundamental principle of detailed balance in quantum statistical mechanics \cite{AccardiEtAl2016ThreeNewPrinciples}. When violated, the system would not reach the correct thermal equilibrium, potentially leading to unphysical results such as negative temperatures or perpetual motion. This makes the KMS condition an essential requirement for any realistic description of open quantum systems in thermal environments.

\subsection{Practical Implementation}
\label{subsec:practical_implementation}

\noindent In practical calculations using the Bloch-Redfield master equation, ensuring that the chosen spectral density satisfies the KMS condition guarantees that the resulting dynamics properly account for the competition between emission and absorption processes. This is crucial for:

\begin{itemize}
    \item Correctly describing thermal equilibration
    \item Modeling temperature-dependent spectroscopic observables
    \item Ensuring physical consistency in long-time behavior
    \item Properly accounting for both spontaneous and thermally-induced transitions
\end{itemize}

\noindent The correlation functions $C_{ij}(t-s)$ defined in Eq. \eqref{eq:Environment_Correlation_Function} are directly related to the spectral density through Fourier transformation, and the KMS condition provides the necessary constraint to ensure physical behavior of the resulting master equation.