% !TEX root = ../main.tex
\chapter{Bath Correlation Functions} % Main chapter title

\label{Chapter_calc_extra_stuff} % Label for referencing this chapter

%----------------------------------------------------------------------------------------
%	SECTION 1
%----------------------------------------------------------------------------------------

\section{Bath Correlator}
\label{sec:bath_correlator}

The general task of this chapter is to calculate the bath correlator, which is defined as:
\begin{equation} \label{eq:bath_correlator}
	C(\tau) = \langle B(\tau) B(0) \rangle,
\end{equation}
where \( B \) is the bath operator, specified later. This correlator measures the operator \( B \) at two different times in the bath.

\section{Useful Tools}

\subsection{Infinite Geometric Series}
An infinite geometric series is expressed as:
\begin{equation} \label{eq:infinite_geometric_series}
	S = a + ar + ar^2 + ar^3 + \dots = \sum_{n=0}^{\infty} ar^n,
\end{equation}
where \(a\) is the first term and \(r\) is the common ratio. For \(|r| < 1\), the sum converges to:
\begin{equation} \label{eq:geometric_series_sum}
	S = \frac{a}{1 - r}.
\end{equation}

Differentiating Eq.~\eqref{eq:geometric_series_sum} with respect to \( r \) yields another useful series:
\begin{equation} \label{eq:derivation_geometric_sum}
	\sum_{n=0}^{\infty} n r^n = \frac{r}{(1-r)^2}, \quad \text{for } |r| < 1.
\end{equation}

%----------------------------------------------------------------------------------------
%	SECTION 2
%----------------------------------------------------------------------------------------

\subsection{Trace}
For a bipartite system \( A \otimes B \), the reduced density matrix of system \( A \) is obtained by tracing out system \( B \):

\begin{equation} \label{eq:partial_trace}
	\rho_A = \mathrm{Tr}_B[\rho_{AB}],
\end{equation}
where the partial trace over the subspace \( B \) is defined as:

\begin{equation} \label{eq:partial_trace_definition}
	\mathrm{Tr}_B\left[\sum_{i,j,k,l} | a_i \rangle \langle a_j | \otimes | b_k \rangle \langle b_l | \right] = \sum_{i,j} | a_i \rangle \langle a_j | \mathrm{Tr}\left[ | b_k \rangle \langle b_l | \right]
\end{equation}

The expectation value of an operator \( A \) in a system \( S \) is calculated as:

\begin{equation} \label{eq:expectation_value}
	\langle A \rangle_S = \mathrm{Tr}_S[\rho_S A] = \frac{1}{Z} \sum_n e^{-\beta E_n} A_{nn},
\end{equation}
where the inverse temperature \(\beta\) is defined as:
\begin{equation} \label{eq:beta_definition}
	\beta = \frac{1}{k_B T}.
\end{equation}

%----------------------------------------------------------------------------------------
%	SECTION 3
%----------------------------------------------------------------------------------------

\section{Harmonic Oscillators}
A bosonic bath is modeled by an infinitely big set of harmonic oscillators, which we assume to be in thermal equilibrium.
For such a system the thermal state is described by the Gibbs distribution:

\begin{equation} \label{eq:gibbs_state}
	\rho = \frac{e^{-\beta H}}{\mathrm{Tr}\left[e^{-\beta H}\right]}, \quad H = \sum_k \hbar \omega_k \left(b_k^{\dagger} b_k + \frac{1}{2}\right), \quad E_k = \hbar \omega_k (n_k + \frac{1}{2}),
\end{equation}
where \( \omega_k \) is the frequency of the \( k \)-th mode, \( b_k^{\dagger} \) and \( b_k \) are the creation and annihilation operators, respectively, and \( n_k = \langle b_k^{\dagger} b_k \rangle \) is the expectation value of the  number operator.
The number operator is an eigenoperator to the number states \(|k\rangle\)
and the number $n_k$ represents the number of excitations in that mode.
The energy of the k-th mode $ E_k $ is directly related to this number.

\subsection{Single Mode}
For a single mode \( k \) harmonic oscillator \( H = \hbar \omega_k b_k^{\dagger} b_k \) in thermal equilibrium at temperature \( T \), the density matrix is given by:

\begin{equation} \label{eq:single_mode_density_matrix}
	\rho = \frac{e^{-\beta \hbar \omega_k b_k^{\dagger} b_k}}{Z_k},
\end{equation}
where \( \omega_k \) is the constant frequency of the mode \( k \).
The partition function \( Z \) can be calculated by the geometric series:

\begin{equation} \label{eq:partition_function}
	Z_k \equiv \mathrm{Tr}\left[e^{-\beta H}\right]
	= \sum_{m=0}^{\infty} \langle m | e^{-\beta \hbar \omega_k (n_k + \frac{1}{2})} | m \rangle
	= e^{-\beta \hbar \omega_k / 2} \sum_{m=0}^{\infty} \langle m | m \rangle \delta_{km} e^{-\beta \hbar \omega_k n_k} \\
	%    = e^{-\beta \frac{\hbar \omega_k }{2}} \sum_{k=0}^{\infty} e^{-\beta \hbar \omega_k k} 
	= \frac{e^{-\beta \hbar \omega_k / 2}}{1 - e^{-\beta \hbar \omega_k}}. % e^{-\beta \frac{\hbar \omega_k}{2}}
\end{equation}
Using this, the expectation value of the number operator can be calculated:

\begin{align} \label{eq:expectation_number_operator}
	n_k & = \langle b_k^{\dagger} b_k \rangle_{\text{th}} = \mathrm{Tr} \left[ b_k^{\dagger} b_k \frac{e^{-\beta H}}{Z_k} \right]                                                             \\
	    & = \frac{\mathrm{Tr} \left[ b_k^{\dagger} b_k e^{-\beta \hbar \omega_k b_k^{\dagger} b_k} \right]}{Z_k}                                                                              \\
	    & = \frac{\sum_{m=0}^{\infty} \langle m|b_k^{\dagger} b_k e^{-\beta \hbar \omega_k b_k^{\dagger} b_k}|m \rangle}{\frac{e^{-\beta \hbar \omega_k / 2}}{1 - e^{-\beta \hbar \omega_k}}} \\
	%    &= \frac{e^{-\beta \hbar \omega_k}}{1 - e^{-\beta \hbar \omega_k}} \\
	    & = \frac{e^{-\beta \hbar \omega_k}}{1 - e^{-\beta \hbar \omega_k}}                                                                                                                   \\
	    & = \frac{1}{e^{\beta \hbar \omega_k} - 1}.
\end{align}
where we have used Eq. \eqref{eq:derivation_geometric_sum} in the last step.
The partition function for the infinite set then generalizes to a product:

\begin{equation} \label{eq:generalized_partition_function}
	Z = \prod_k Z_k = \prod_k \frac{e^{-\beta \hbar \omega_k / 2}}{1 - e^{-\beta \hbar \omega_k}}.
\end{equation}
%----------------------------------------------------------------------------------------
%	SECTION 4
%----------------------------------------------------------------------------------------

\section{Bath Correlators}
\label{sec:bath_corr_trans_rates}

Now we turn to calculating the bath correlator.
The bath operator \( B \) is defined as:

\begin{equation} \label{eq:bath_operator}
	B = \sum_{n=1}^{\infty} c_n x_n,
\end{equation}
where \( c_n \) are coupling constants and \( x_n \) are position operators. Note that in Sec.~\ref{sec:Derivation_redfield_eq_from_microscopic_dynamics}, the interaction Hamiltonian reduces such that only one bath correlator \( C(\tau) \equiv C_{ii}(\tau) \) in Eq.~\eqref{eq:Environment_Correlation_Function} remains.

Expressing \( B \) in terms of creation and annihilation operators:

\begin{align}
	B(0)    & = \sum_{n=1}^{\infty} c_n \sqrt{\frac{1}{2 m_n \omega_n}} (b_n + b_n^\dagger), \label{eq:bath_operator_t0}                                                        \\
	B(\tau) & = \sum_{n=1}^{\infty} c_n \sqrt{\frac{1}{2 m_n \omega_n}} \left( b_n e^{-i \omega_n \tau} + b_n^\dagger e^{i \omega_n \tau} \right). \label{eq:bath_operator_tau}
\end{align}

Substituting \( B(\tau) \) and \( B(0) \) into Eq.~\eqref{eq:bath_correlator}, we find:

\begin{equation} \label{eq:correlator_substitution}
	C(\tau) = \left\langle \sum_{n=1}^{\infty} c_n \sqrt{\frac{1}{2 m_n \omega_n}} (b_n e^{-i \omega_n \tau} + b_n^\dagger e^{i \omega_n \tau}) \sum_{m=1}^{\infty} c_m \sqrt{\frac{1}{2 m_m \omega_m}} (b_m + b_m^\dagger) \right\rangle.
\end{equation}

Similar calculations to Eq. \eqref{eq:expectation_number_operator} thermal expectation values can be calculated:

\begin{align} \label{eq:thermal_expectations}
	\langle b_n b_m^\dagger \rangle & = \delta_{nm} (n_n + 1), \quad \langle b_n^\dagger b_m \rangle = \delta_{nm} n_n,
\end{align}
where \( n_k \) is the Bose-Einstein distribution:

\begin{equation} \label{eq:bose_einstein_distribution}
	n_k = \frac{1}{e^{\beta \omega_k} - 1},
\end{equation}
we obtain:

\begin{equation} \label{eq:correlator_result}
	C(\tau) = \sum_{n=1}^{\infty} \frac{c_n^2}{2 m_n \omega_n} \left[ (n_n + 1) e^{-i \omega_n \tau} + n_n e^{i \omega_n \tau} \right].
\end{equation}

\subsection{Spectral Density Representation}
\label{subsec:spectral_density}

The bath correlator can be expressed in terms of the spectral density \( J(\omega) \), defined as:
\begin{equation} \label{eq:spectral_density}
	J(\omega) = \pi \sum_{n=1}^{\infty} \frac{c_n^2}{2 m_n \omega_n} \delta(\omega - \omega_n).
\end{equation}

With this, the bath correlator becomes:
\begin{equation} \label{eq:correlator_spectral_density}
	C(\tau) = \int_0^\infty d\omega \frac{J(\omega)}{\pi} \left[ (n(\omega) + 1) e^{-i \omega \tau} + n(\omega) e^{i \omega \tau} \right].
\end{equation}

\subsection{Continuum Limit}
As the frequencies \( \{\omega_j\} \) become dense, the sum transitions to an integral:
\[
	\sum_{j=1}^{\infty} \quad \longrightarrow \quad \int d\omega'\,\rho(\omega'),
\]
where \( \rho(\omega') \) is the density of states. The coupling \( g_j \) becomes a function of frequency, \( g(\omega') \), leading to:
\[
	J(\omega) = \rho(\omega)\,g(\omega)^2.
\]

In this continuum limit, the bath correlator Eq. \eqref{eq:correlator_result} becomes:
\begin{equation} \label{eq:correlator_final}
	C(\tau) = \int_0^\infty d\omega \frac{J(\omega)}{\pi} \left[ \coth\left( \frac{\beta \omega}{2} \right) \cos(\omega \tau) - i \sin(\omega \tau) \right].
\end{equation}

\subsection{Ohmic Spectral Density}
\label{subsec:ohmic_spectral_density}

Source: Ulrich Weiss chapter 7.3:
(*my interpretation: The Redfield equation is influenced by classical phenomenological models of dissipation*)
For an Ohmic spectral density, the damping is frequency-independent, and the spectral density is given by:
\begin{equation} \label{eq:ohmic_spectral_density}
	J(\omega) \propto \gamma \omega,
\end{equation}
where \( \gamma \) is the damping constant. To ensure physical behavior, a cutoff is introduced:
\begin{equation} \label{eq:cutoff_spectral_density}
	J(\omega) = \eta \frac{\omega^s}{\omega_c^{s-1}} e^{-\omega / \omega_c},
\end{equation}
where \( \eta \) is a dimensionless coupling constant, \( \omega_c \) is the cutoff frequency, and \( s \) determines the type of spectral density (Ohmic for \( s = 1 \), sub-Ohmic for \( s < 1 \), and super-Ohmic for \( s > 1 \)).
